AI Assistance Declaration: I used ChatGPT (GPT-5.6 Sol) for brainstorming documentation structure, Markdown formatting suggestions, wording refinement, and review support. Prompts used are listed in Appendix.md. I verified the final content by reviewing the Markdown structure, checking the R Markdown organization, and confirming that no real or personal data is used. All final calculations are done by myself. I am responsible for the accuracy and originality of this work.

---
title: "Sales Summary Script Documentation"
author: "Douha Suliman"
date: "September 25, 2026"
output: html_document
---

# Purpose

This R Markdown document explains the purpose, inputs, processing steps, and outputs of the `sales_summary.R` script.

The script demonstrates a simple workflow using a **synthetic sales dataset**. It imports the data, calculates revenue for each row, summarizes revenue by product, and prints the final summary.

# Inputs

The script uses one input file:

- `data/synthetic_sales.csv`

The synthetic dataset contains the following fields:

- `OrderID` — synthetic order identifier.
- `Product` — product name.
- `Quantity` — number of units.
- `UnitPrice` — price per unit.

# Script Logic

## 1. Import the Dataset

The script reads the CSV file with base R:

```{r eval=FALSE}
sales <- read.csv("data/synthetic_sales.csv")
```

## 2. Create a Revenue Field

Revenue is calculated for each row by multiplying quantity by unit price:

```{r eval=FALSE}
sales$Revenue <- sales$Quantity * sales$UnitPrice
```

## 3. Summarize Revenue by Product

The `aggregate()` function groups the data by product and sums the revenue values:

```{r eval=FALSE}
revenue_by_product <- aggregate(
  Revenue ~ Product,
  data = sales,
  FUN = sum
)
```

## 4. Display the Result

The final summary is printed in the R console:

```{r eval=FALSE}
print(revenue_by_product)
```

# Outputs

The script produces a table with:

- one row per product;
- a product name; and
- total revenue for that product.

# Validation

I reviewed the document for:

- consistent Markdown and R Markdown headings;
- correctly fenced R code chunks;
- matching file names and paths;
- clear descriptions of inputs and outputs; and
- use of synthetic rather than personal or real data.

The code chunks are shown with `eval=FALSE` so the document focuses on explaining the script rather than using AI-generated calculations.

# Conclusion

This file provides a concise explanation of a single R script and demonstrates how R Markdown can be used to create clear technical documentation.
